Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-05-06 paper/paper.tex
saturday pass

\section{Introduction}
%
In the absence of external stimulation, the human brain is not at rest.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that probably maintains
baseline mental activities
\citep{raichle2001pnas, randy2008, bzdok2015resting}.
The ``default mode network'' (DMN) continues
to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
during task-unconstrained behavior
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is only weakly modulated
at the onset of defined psychological tasks \citep{raichle_baseline}.
At the opposite, during sleep, the
decoupling of brain structures discarded the idea of the DMN being
only a passive network resonance and rather supported an
important role in sustaining conscious awareness \citep{horovitz2009decoupling}.

@@

\subsection{Towards a formal account of default mode function: higher-order control of the organism}
The human DMN is responsible for a large part of the baseline neural activity,
which typically decreases when engaged in psychological experiments
\citep{raichle_baseline}.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
Today, many psychologists and neuroscientists believe that
the DMN implements some form of
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008, spreng2009common}.
This brain network
might have emerged to continuously predict the environment using
mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode regions \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into DMN function is
particularly challenging because
this network appears to
simultaneously impact perception-action cycles and
maintain probabilistic contemplations
across time, space, and content domains
\citep{boyer2008evolutionary}.

The present work adopts the perspective of
a human agent faced with the choice of the next actions
and guided by outcomes
of really happened, hypothetically imagined, and
expected futures to optimize behavioral performance.
Formally, a particularly attractive framework
to describe, quantify, and predict intelligent systems, such as the brain,
is proposed to be the combination of control theory and
reinforcement learning (RL).
An intelligent agent improves its interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it is known that, the more the world is easy to predict,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Without requiring explicit awareness by the agent,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that can be performed
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of really performed action,
the DMN dynamics could be constantly adapted in feedback loops
shaped by prediction error.
Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.

We argue that DMN implication in many
advanced cognitive processes in humans
can be recast as prediction error minimization
based on probabilistic mental simulations,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide blocks of pseudo-experience to iteratively
optimize the behavior of the organism.
%
It has indeed been proposed that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea is invigorated by
empirical evidence from
neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

@@

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
%
This DMN node has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment
\citep{amygdala_bzdok}.

The putative role of the PMC in continuous information integration of
environmental relevance
and ensuing top-level guidance of action on the environment is supported
by many neuroscience experiments.
Electrophysiological recordings in animals implicated PMC neurons in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
outcome-dependent behavioral modulation \citep{hayden2009electrophysiological},
as well as approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference for or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its actual value,
based on an ``internal currency of value'' \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons
promoted exploratory actions,
which would otherwise be shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from neuronal spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different parts of auditory narratives in an fMRI study
\citep{simony2016dynamic}.

Moreover, the retrosplenial portion of the PMC can subserve
representation of action possibilities
and evaluation of reward outcomes by integration with
information from memory and different perspective frames.
Regarding memory recall, retrosplenial damage have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in rabbits and humans
\citep{vann2009does}.
Regarding perspective frames, the retrosplenial subregion of the PMC has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory environment) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible actions and integrating that information
with memory and perspective frames
into short- and longer-term behavioral agendas.
Estimated value, that differs across individuals, probably enriches
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt the organism to changes
in both the external environment and its internal representation
to enable strategic behavior.

@@

The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
errors in these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
construction of probabilistic internal information drawing from
episodic memory recall, mental scene elaboration processes,
and stored knowledge of the world.
%
Moreover,
in a computational neuroimaging experiment,
dorsomedial PFC activity preferentially processed the consequences of
action choices that were considered but not actually executed,
whereas ventromedial PFC (vmPFC) activity
processed especially value outcomes of performed actions in the environment
\citep{nicolle2012agent}.

Comparing to the dmPFC,
the vmPFC probably subserves
subjective value evaluation and
risk estimation of relevant environmental stimuli.
This DMN node may
orchestrate adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{rolls_OFC, doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based action choice
\citep{glascher2012lesion}.
Indeed,
this DMN node is preferentially connected with reward-related and limbic regions.
The vmPFC is well known to have strong axonal connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals
from dopaminergic neurotransmitter pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
subserve reward processing, and thus perhaps RL,
in humans more broadly \citep{doherty2015structure}.
Consistently, diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative}
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two different functional connectivity analyses in humans also strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC \citep{bzdok2015subspecialization}.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to represent triggered
emotional and motivational states \citep{damasio1996somatic}.
Such real or imagined arousal states could be mapped in the vmPFC
as a bioregulatory disposition influencing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
value estimates
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in the vmPFC \citep{Schuck2016}.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from the sensory environment.

@@

While the HC
is traditionally believed to allow recalling past experience,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017,
boyer2008evolutionary}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in re-experiencing the past (i.e., amnesia),
but also imagining one's own future and
imagined experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by patients with HC lesion exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal HC revealed
some constantly active neuronal populations whose firing coincided with
specific locations in space during environmental navigation.
Indeed, when an animal is choosing between alternative
paths, the corresponding neuronal populations in the HC
spike one after another  \citep{johnson2007neural}.
Such neuronal patterns in the HC appear to directly indicate upcoming behavior,
such as in planning navigational trajectories
\citep{pfeiffer2013hippocampal} and
memory consolidation of choice relevance \citep{lavilleon2015}.
Congruently,
London taxi drivers, humans with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
HC \citep{maguire2000navigation}.

There is increasing evidence that
HC function extends beyond simple forms of
encoding and reconstruction of memory and spatial information.
Based on neuronal spike recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited by experimental task design,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, neuronal spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such spking patterns in hippocampal neurons during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
Moreover, in mice, invasively triggering spatial experience recall
in the HC during sleep
has been shown to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
probably contribute to advanced cognitive processes that require
re-experiencing or newly constructed mental scenarios,
such as in recalling autobiographical memory episodes
\citep{hassabis2007patients}.

Thus, the HC probably orchestrates re-experience of environmental aspects for
consolidations based on reenactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may impact
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

\subsection{The right and left TPJ: prediction error signaling and world semantics}
The DMN emerges with its midline structures early in human development
\citep{doria2010}, while
the right and left TPJ may become fully integrated into this major brain
network only after birth.
The TPJs are known to exhibit hemispheric differences
based on microanatomical properties and gyrification pattern
\citep{seghier2013angular}.
Globally, neuroscientific investigations on hemispheric functional specialization
have highlighted the right versus left cerebral hemisphere as dominant for
attentional versus semantic functions
\citep{seghier2013angular, bzdok2013tpj, bzdok2016left}.

The TPJ in the right-hemispheric DMN (RTPJ)
has been shown to be closely related to
multi-sensory prediction and prediction error signaling.
It is probably central for
action initiation during goal-dircted psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Involvement of this DMN node was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005}, and
detection of environmental changes across
visual, auditory, or tactile stimulation in an fMRI experiment
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and ensuing outcomes,
without necessarily relating those to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Additionally, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, patients with RTPJ damage have particular difficulties
with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on self-relevant mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
The RTPJ might hence be an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with (unexpected) environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

>>>>

\section{Introduction}
%
In the absence of external stimulation, the human brain is not at rest.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that probably maintains
baseline mental activities
\citep{raichle2001pnas, randy2008, bzdok2015resting}.
The ``default mode network'' (DMN) continues
to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
during free-ranging thought
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is only weakly modulated
at the onset of defined psychological tasks \citep{raichle_baseline}.
At the opposite, during sleep, the
decoupling of brain structures discarded the idea of the DMN being
only a passive network resonance and rather supported an
important role in sustaining conscious awareness \citep{horovitz2009decoupling}.

@@

\subsection{Towards a formal account of default mode function: higher-order control of the organism}
The human DMN is responsible for a large part of the baseline neural activity,
which typically decreases when engaged in psychological experiments
\citep{raichle_baseline}.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
Today, many psychologists and neuroscientists believe that
the DMN implements some form of
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008, spreng2009common}.
This brain network
might have emerged to continuously predict the environment using
mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode regions \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into DMN function is
particularly challenging because
this network appears to
simultaneously modulate perception-action cycles and
maintain probabilistic contemplations
across time, space, and content domains
\citep{boyer2008evolutionary}.

The present work adopts the perspective of
a human \textit{agent} faced with the choice of the next actions
and guided by outcomes
of really happened, hypothetically imagined, and
expected futures to optimize behavioral performance.
Formally, a particularly attractive framework
to describe, quantify, and predict intelligent systems, such as the brain,
is proposed to be the combination of control theory and
reinforcement learning (RL).
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it is known that, the more the world is easy to predict,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly adapted in feedback loops
shaped by prediction error.
Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.

We argue that DMN implication in many
advanced cognitive processes in humans
can be recast as prediction error minimization
based on probabilistic mental simulations,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide blocks of pseudo-experience to iteratively
optimize the behavior of the organism.
%
It has indeed been proposed that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

@@

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN node has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok}.

The putative role of the PMC in continuous abstract integration of
environmental relevance
and ensuing top-level guidance of action on the environment is supported
by many neuroscience experiments.
Electrophysiological recordings in animals implicated PMC neurons in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
outcome-dependent behavioral modulation \citep{hayden2009electrophysiological},
as well as approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would pursue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference for or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its actual value,
based on an ``internal currency of value'' \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons
promoted exploratory actions,
which would otherwise be shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changes in upcoming choice trials, while their neural patterns were
distinct from neuronal spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different parts of auditory narratives in an fMRI study
\citep{simony2016dynamic}.

Moreover, the retrosplenial portion of the PMC could support
representation of action possibilities
and evaluation of reward outcomes by integrating
information from memory recall and different perspective frames.
Regarding memory recall, retrosplenial damage have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in animals and humans
\citep{vann2009does}.
Regarding perspective frames, the retrosplenial subregion of the PMC has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory environment) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible actions and integrating that information
with memory and perspective frames
into short- and longer-term behavioral agendas.
Estimated value, that differs across individuals, probably enriches
statistical assessment of the environment
to map and predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt the organism to changes
in both the external environment and its internal representation
to enable strategic behavior.

@@

The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
construction of probabilistic internal information drawing from
memory recall, mental scene elaboration processes,
and stored knowledge of world regularities.
%
Moreover,
in a computational neuroimaging experiment,
dorsomedial PFC activity preferentially processed the consequences of
action choices that were considered but not actually executed,
whereas ventromedial PFC (vmPFC) activity
processed especially value outcomes of actions performed on the environment
\citep{nicolle2012agent}.

Comparing to the dmPFC,
the vmPFC probably subserves
subjective value evaluation and
risk estimation of relevant environmental stimuli.
The ventromedial prefrontal DMN may
orchestrate adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{rolls_OFC, doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based action choice
\citep{glascher2012lesion}.
Indeed,
this DMN node is preferentially connected with reward-related and limbic regions.
The vmPFC is well known to have strong axonal connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals
from dopaminergic neurotransmitter pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
subserve reward processing, and thus perhaps RL,
in humans more broadly \citep{doherty2015structure}.
Consistently, diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative}
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two different functional connectivity analyses in humans also strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC \citep{bzdok2015subspecialization}.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to represent triggered
emotional and motivational states \citep{damasio1996somatic}.
Such real or imagined arousal states could be mapped in the vmPFC
as a bioregulatory disposition influencing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
value estimates
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in the vmPFC \citep{Schuck2016}.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from the sensory environment.

@@

While the HC
is traditionally believed to allow recalling past experience,
there is now increasing evidence for an important role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017,
boyer2008evolutionary}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in re-experiencing the past (i.e., amnesia),
but also imagination dedicated to one's own future and
imagination of experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by patients with HC lesion exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal HC revealed
some constantly active neuronal populations whose firing coincided with
specific locations in space during environmental navigation.
Indeed, when an animal is choosing between alternative
paths, the corresponding neuronal populations in the HC
spike one after another  \citep{johnson2007neural}.
Such neuronal patterns in the HC appear to directly indicate upcoming behavior,
such as in planning navigational trajectories
\citep{pfeiffer2013hippocampal} and
memory consolidation of choice relevance \citep{lavilleon2015}.
Congruently,
London taxi drivers, humans with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
HC \citep{maguire2000navigation}.

There is increasing evidence that
HC function extends beyond simple forms of
encoding and reconstruction of memory and space information.
Based on spike recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited by experimental task design,
have been shown to be re-enacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, neuronal spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Similar spiking patterns in hippocampal neurons during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
Moreover, in mice, invasively triggering spatial experience recall
in the HC during sleep
has been shown to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
probably contribute to advanced cognitive processes that require
re-experiencing or newly constructed mental scenarios,
such as in recalling autobiographical memory episodes
\citep{hassabis2007patients}.

Thus, the HC probably orchestrates re-experience of environmental aspects for
consolidations based on re-enactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
As such, the HC may impact
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

\subsection{The right and left TPJ: prediction error signaling and world semantics}
The DMN emerges with its midline structures early in human development
\citep{doria2010}, while
the right and left TPJs may become fully integrated into this major brain
network only after birth.
The TPJs are known to exhibit hemispheric differences
based on microanatomical properties and gyrification pattern
\citep{seghier2013angular}.
Globally, neuroscientific investigations on hemispheric functional specialization
have highlighted the right versus left cerebral hemisphere as dominant for
attentional versus semantic functions
\citep{seghier2013angular, bzdok2013tpj, bzdok2016left}.

The TPJ in the right-hemispheric DMN (RTPJ)
has been shown to be closely related to
multi-sensory prediction and prediction error signaling.
It is probably central for
action initiation during goal-dircted psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Involvement of this DMN node was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005}, and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and ensuing outcomes,
without necessarily relating those to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Additionally, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, patients with RTPJ damage have particular difficulties
with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
The RTPJ might hence be an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The TPJ in the left-hemispheric DMN (LTPJ),
in turn, has a close relationship to Wernicke's area
involved in semantic processes, such as in spoken and written language.
Neurological patients with damage in Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Patient speech
preserves natural rhythm and normal syntax, yet the
voiced sentences lack meaning (i.e., aphasia).
Abstracting from speech interpretations in linguistics
and neuropsychology,
the LTPJ probably mediates access to and integration of world knowlege,
such as required during action considerations
\citep{binder2011neurobiology, seghier2013angular}.
For instance, LTPJ damage in humans also entails problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing explicit language content
\citep{varney1987locus}.
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Inner speech also hinges on knowledge recall
about the physical and social world.
Indeed,
the internal production of
verbalized thought (``language of the mind'') was closely related to the LTPJ
in a pattern analysis of brain volume
\citep{geva2011neural}.
Further,
episodic memory recall and mental imagery strongly draw on
reassembling world knowledge.
Isolated building blocks of world structure probably get rebuilt
in internally constructed mental scenarios that
guide present action choice,
weigh hypothetical possibilities, and forecast future events.
%
The LTPJ may hence facilitate the automated environmental predictions
by incorporating experience-derived building blocks of world regularities
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation
